Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-25 robdml.tex
tex

Then, how hard should such ``\emph{destination} hardness'' be?
%
Recall that the model is already prone to collapse with excessively hard benign
triplets, let alone adversarial examples.
%
Thus, intuitively, the \emph{destination} hardness can be that of another
benign triplet which is harder than the given triplet (\eg, a
Semihard~\cite{facenet} triplet).
%
However, in the late phase of training, the
expectation of \emph{destination} hardness of another benign triplet
will be small, and close to the \emph{hardness} expectation of the given triplet.
%
This leads to weak attacks, and inefficient learning from a weak adversary.
%
Besides, we speculate a strong adversary in the early phase of DML training may
hinder the model from learning good embeddings, and hence influence the
eventual performance on benign examples.
%
Based on these, a \emph{destination} hardness calculated from benign examples
may not be the best choice.
%
A better \emph{destination} hardness should be able to balance the training
objectives in the early and late phases of training.

@@

@@

In the traditional min-max adversarial training setting~\cite{madry}, the inner
``max'' problem should maximize $H(X_a,X_p^S,X_n^S)$ and hence maximize the triplet
loss.
%
However, existing defense methods~\cite{advorder,robrank} refrain from adopting
such paradigm because the model will quickly collapse with excessively hard
triplets~\cite{facenet}, and adopt relatively weak adversaries instead
(inefficient in gaining robustness).

Then, we want change (increase) the hardness level from $H_\mathsf{S}$ into $H_\mathsf{D}$, by
finding adversarial examples for $(X_a, X_p^S, X_n^S)$.
%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
We denote the the hardness of adversarial examples as $\tilde{H}_S=H(X_a{+}r_a,
X_p^S{+}r_p, X_n^S{+}r_n)$.
%
This is named as \emph{Hardness Manipulation} (HM), which is implemented as
follows:
%
\begin{equation}
	%
	\bar{r}_a, \bar{r}_p, \bar{r}_n = \argmin_{r_a,r_p,r_n} \big\|\min(0,
	\tilde{H}_\mathsf{S} - H_\mathsf{D}) \big\|_2^2.
	%
	\label{eq:hm}
	%
\end{equation}

@@

>>>>

Then, how hard should such ``\emph{destination} hardness'' be?
%
Recall that the model is already prone to collapse with excessively hard benign
triplets~\cite{facenet}, let alone adversarial examples.
%
Thus, intuitively, the \emph{destination} hardness can be that of another
benign triplet which is harder than the given triplet (\eg, a
Semihard~\cite{facenet} triplet).
%
However, in the late phase of training, the
expectation of \emph{destination} hardness of another benign triplet
will be small, and close to the \emph{hardness} expectation of the given triplet.
%
This leads to weak attacks, and inefficient learning from a weak adversary.
%
Besides, we speculate a strong adversary in the early phase of DML training may
hinder the model from learning good embeddings, and hence influence the
eventual performance on benign examples.
%
Based on these, a \emph{destination} hardness calculated from benign examples
may not be the best choice.
%
A better \emph{destination} hardness should be able to balance the training
objectives in the early and late phases of training.

@@

% define hardness

@@

% define hardness manipulation

Then, \emph{Hardness Manipulation} (HM) aims to increase the \emph{source} hardness
$H_\mathsf{S}$ into a specified ``\emph{destination} hardness'' $H_\mathsf{D}$,
by finding adversarial examples of the source triplet $(\mA, \mP, \mN)$, \ie,
$(\mA + \hat{\vr}_a, \mP + \hat{\vr}_p, \mN + \hat{\vr}_n)$,
where $(\hat{\vr}_a, \hat{\vr}_p, \hat{\vr}_n)$ are the adversarial perturbations.
%
If we denote the hardness of the adversarially perturbed \emph{source} triplet
as $\tilde{H}_\mathsf{S}$, the HM can be implemented as follows:
%
\begin{equation}
	%
	\hat{\vr}_a, \hat{\vr}_p, \hat{\vr}_n = \argmin_{r_a,r_p,r_n} \big\|\max(0,
	H_\mathsf{D} - \tilde{H}_\mathsf{S} ) \big\|_2^2.
	%
	\label{eq:hm}
	%
\end{equation}

%
Note, it is not expected to lower the hardness during the attack, namely
$E[H_\mathsf{D}]$ should be greater or equal to $E[H_\mathsf{S}]$.
%
%

@@

In the traditional min-max adversarial training setting~\cite{madry}, the inner
``max'' problem should maximize $H(X_a,X_p^S,X_n^S)$ and hence maximize the triplet
loss.
